Supernovae Time Profiles as a Portal to New Physics at Neutrino Telescopes

### 2024-01-23 paper/sections/preamble.tex

\begin{abstract}
Neutrino telescopes, such as IceCube, have promised precise time resolutions at the millisecond level for detecting collective rise signals in all photomultiplier.
We demonstrate that such precise time resolutions will play a crucial role in tracking neutrino signals with novel timing patterns related to weakly interacting massive particles in physics beyond the Standard Model.
Focusing on future Supernova explosion in the operation time of IceCube, Focusing on a future Supernova explosion in the operation time of IceCube as an example, we show that neutrino telescopes can provide unprecedented probes of new particles with both low and high masses, where neutrino signals are advanced and delayed, respectively, while comparable to DUNE and Hyper-K in the intermediate mass region.
\end{abstract}

>>>>

\begin{abstract}
Neutrino telescopes, such as IceCube, have promised precise time resolutions at the millisecond level for detecting collective rise signals in all photomultiplier.
We demonstrate that such precise time resolutions will play a crucial role in tracking neutrino signals with novel timing patterns related to weakly interacting massive particles in physics beyond the Standard Model.
Focusing on a future Supernova explosion in the operation time of IceCube as an example, we show that neutrino telescopes can provide unprecedented probes of new particles with both low and high masses from probing early-glow and after-glow neutrino signals, respectively, while comparable to DUNE and Hyper-K in the intermediate mass region.
\end{abstract}

### 2024-01-23 paper/sections/supplements.tex

\section{Appendix A: Dipole Magnetic Moment Portal}
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/magnetic_moment_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the magnetic moment case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, and two next-generation neutrino experiments, Hyper-Kamiokande and DUNE. The shaded regions are excluded by the energy loss constraints, non-observation of photon and neutrino signals from SN1987A studied in \cite{Brdar:2023tmi} and by   constraints on the energy release from SN explosion following \cite{PhysRevLett.128.221103}.
    }
    \label{fig:magnetic_moment_sensitivity}
\end{figure}
The first scenario under consideration is active-to-sterile neutrino transition magnetic moment described by \cite{Magill:2018jla,Brdar:2020quo,Brdar:2023tmi}
\begin{align}
    \mathcal{L} \supset \sum_\alpha d_\alpha \bar{N}\sigma_{\mu\nu} \nu^{\alpha} F^{\mu\nu}-\frac{M_N}{2} \bar{N}^c N + \text{h.c.}\,,
    \label{eq:Lag}
\end{align}
where $\nu^{\alpha}$ and $N$ represent active and sterile neutrinos, respectively. Further, $F^{\mu\nu}$ is the field strength tensor of the electromagnetic field and $d_\alpha$ is the dimensionful coefficient of this dimension-5 term and $M_N$ is sterile neutrino mass. We will abbreviate $d_\alpha \equiv d$ given that we assume flavor universal interaction. The dominant production channels for $N$ inside the SN are $\nu e^- \to \nu e^-$ at lower energies and $\nu \gamma \to N$ for larger active neutrino energies. Both processes occur due to the interaction term in \cref{eq:Lag}; after $N$ are produced, they decay to active neutrinos and photons which is again realized through the same term in the Lagrangian, with the decay width for $N\to\nu\gamma$ given by $\Gamma_N = 6d^2 M_N^3/4 \pi$ \cite{Plestid:2020vqf}. We notice that the time delay near the exclusion boundary are typically of $\Delta t \sim \mathcal{O}(1)$sec, which overlaps with the standard burst time scale. We don't expect improvements over DUNE and Hyper-K.

{\color{red} Check the bg for DUNE and Hyper-K is the atmospheric neutrino bg which should also contribute to our analysis? IceCube sees at}

>>>>

\section{Appendix A: Dipole Magnetic Moment Portal}
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/magnetic_moment_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the magnetic moment case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, and two next-generation neutrino experiments, Hyper-Kamiokande and DUNE. The shaded regions are excluded by the energy loss constraints, non-observation of photon and neutrino signals from SN1987A studied in \cite{Brdar:2023tmi} and by constraints on the energy release from SN explosion following \cite{PhysRevLett.128.221103}.
    }
    \label{fig:magnetic_moment_sensitivity}
\end{figure}
The first scenario under consideration is active-to-sterile neutrino transition magnetic moment described by \cite{Magill:2018jla,Brdar:2020quo,Brdar:2023tmi}
\begin{align}
    \mathcal{L} \supset \sum_\alpha d_\alpha \bar{N}\sigma_{\mu\nu} \nu^{\alpha} F^{\mu\nu}-\frac{M_N}{2} \bar{N}^c N + \text{h.c.}\,,
    \label{eq:Lag}
\end{align}
where $\nu^{\alpha}$ and $N$ represent active and sterile neutrinos, respectively. Further, $F^{\mu\nu}$ is the field strength tensor of the electromagnetic field and $d_\alpha$ is the dimensionful coefficient of this dimension-5 term and $M_N$ is sterile neutrino mass. We will abbreviate $d_\alpha \equiv d$ given that we assume flavor universal interaction. The dominant production channels for $N$ inside the SN are $\nu e^- \to N e^-$ at lower energies and $\nu \gamma \to N$ for larger active neutrino energies. Both processes occur due to the interaction term in \cref{eq:Lag}; after $N$ are produced, they decay to active neutrinos and photons which is again realized through the same term in the Lagrangian, with the decay width for $N\to\nu\gamma$ given by $\Gamma_N = 6d^2 M_N^3/4 \pi$ \cite{Plestid:2020vqf}. The limits from IceCube are shown in \cref{fig:magnetic_moment_sensitivity}, together with previously excluded regions by energy loss, non-observations of photons and neutrinos from SN1987A \cite{Brdar:2023tmi}. We also estimates the excluded regions (xxx shaded) where photons from $N$ decay inside the photosphere exceeds the energy release observed following \cite{PhysRevLett.128.221103}.
We notice that the time delay near the exclusion boundary are typically of $\Delta t \sim \mathcal{O}(1)$ sec, which overlaps with the standard burst time scale. We don't expect improvements over DUNE and Hyper-K. We also point out by neglecting the $N$ production channel $\nu p^+\to N p^+$ inside supernove, all the limits presented in \cref{fig:magnetic_moment_sensitivity} are conservative.
